Sparse Function-space Representation of Neural Networks

### 2023-09-04 workshops/icml-duality/main.tex
commit after getting paper into good state

% Short section names etc
% This must be imported last!
%\usepackage{cleveref}
\usepackage[capitalise,nameinlink]{cleveref}
\crefname{section}{Sec.}{Secs.}
\crefname{algorithm}{Alg.}{Algs.}
\crefname{appendix}{App.}{Apps.}
\crefname{definition}{Def.}{Defs.}
\crefname{table}{Table}{Tables}

@@

We consider supervised learning with inputs $\vx_i \in \inputDomain$ and outputs $\vy_i \in \outputDomain$ (e.g. regression) or $\vy_{i} \in \{0,1\}^{C}$ (e.g. classification),
giving a data set $\dataset = \{(\vx_{i} , \vy_{i})\}_{i=1}^{N}$.
We introduce a  NN $f_\mathbf{w}: \inputDomain \to \outputDomain$ with weights $\weights \in \R^{P}$ and use a likelihood function $p(\vy \mid f_\mathbf{w}(\vx))$
to link the function values to the output $\vy$ (Gaussian for regresssion or Categorical for classification).
For notational conciseness we stick to scalar outputs $y_{i}$.

@@

\textbf{Laplace-GGN}
The Laplace approximation \citet{aa} builds upon this observation and approximates the weight-space posterior $p(\vw \mid \mathcal{D}) = \mathcal{N}(\vw^{*} , \bm\Sigma)$
around the MAP weights ($\vw^{*}$) by setting the covariance to the Hessian of the posterior,
\begin{align} \label{eq:laplace-cov}
 \bm\Sigma = - \left[\nabla^{2}_{\vw\vw} \log p(\vw \mid \mathcal{D})  |_{\vw=\vw^{*}} \right]^{-1}.
\end{align}
Computing \cref{eq:laplace-cov} requires calculating the Hessian of the log-likelihood $\nabla^{2}_{\vw\vw} \log p(\vy \mid f_{\vw}(\vx))$ from \cref{eq-empirical-risk}.
As highlighted in \citet{immer2021improving}, computing this Hessian is often infeasible and in practice it is common to adopt the
generalised Gauss-Newton (GGN) approximation,
\begin{align}
 \nabla^{2}_{\vw\vw} \log p(\vy \mid f_{\vw}(\vx)) \approx \Jac{\vw}{\vx}^{\top} \nabla_{\vf\vf}^{2}\log(\vy\mid\vf) \Jac{\vw}{\vx}.
\end{align}
% \textbf{Laplace-GGN converts BNN into GLM}
\citet{immer2021improving} highlighted that the GGN approximation corresponds to a local linearization of the NN,
% Recent work \citep{khan2019approximate,maddox2021fast} has converted NNs to GPs by linearizing around the NN's MAP weights (obtained from \cref{eq-empirical-risk}).
% Recent work \citep{khan2019approximate,maddox2021fast} has shown that NN (in weight space) leads to a GP, i.e. a function-space approximation.
$f_{\weights^{*}}^{\text{lin}}(\vx) = f_{\weights}(\vx) + \Jac{\weights^{*}}{\vx}(\weights-\weights^{*})$.
This suggests that predictions should be made with a generalised linear model (GLM), given by,
\begin{align} \label{eq-glm}
  p_{\text{\sc GLM}}(\vy \mid \vx, \mathcal{D}) &= \E_{q(\vw)} \left[ p(\vy \mid f_{\vw^{*}}^{\text{lin}}(\vx))\right]. %\nonumber \\
  % &\approx \frac{1}{S} \sum_{s} p(\vy \mid f_{\vw_{s}}^{\text{lin}}(\vx)), \quad \vw_{s} \sim q(\vw)
\end{align}

\textbf{Weight space GLM is equivalent to a GP}
As Gaussian distributions remain tractable under linear transformations, a linear function in terms of parameters can be converted from the weight space to the function space \citep[see Ch.~2.1 in ][]{rasmussen2006gaussian}.
 % highlighted that the GGN approximation corresponds to a local linearization of the NN,
As such (and show in \citet{immer2021improving}), the Bayesian GLM in \cref{eq-glm} has an equivalent GP formulation,
% \begin{align}
%   p_{\text{\sc GP}}(\vy \mid \vx, \mathcal{D}) &= \E_{q(\vf)} \left[ p(\vy \mid \vf)\right] %\nonumber \\
%   % &\approx \frac{1}{S} \sum_{s} p(\vy \mid f_{\vw_{s}}^{\text{lin}}(\vx)), \quad \vw_{s} \sim q(\vw)
% \end{align}
% where the GP posterior is given by,
\begin{subequations}  \label{eq-gp-pred-immer}
\begin{align}
  p_{\text{\sc GP}}(\vy \mid \vx, \mathcal{D}) &= \E_{q(\vf)} \left[ p(\vy \mid \vf)\right] \\ %\nonumber \\
  \myexpect_{p(f_i \mid\vy)}[f_i] &= f_{\vw^{*}}(\vx_{i}), \\
  \mathrm{Var}_{p(f_i \mid \vy)}[f_i] &= k_{ii} - \vk_{\vx i}^\top ( \MKxx + \bm\Lambda^{-1})^{-1} \vk_{\vx i},
\end{align}
\end{subequations}
where the kernel $\kappa(\vx,\vx')=\frac{1}{\delta} \Jac{\weights^*}{\vx} \, \JacT{\weights^*}{\vx'}$ is the Neural Tangent Kernel \citep[NTK,][]{jacot2018neural},
based on the Jacobian at the MAP $\Jac{\weights}{\vx} \coloneqq \left[ \nabla_\weights f_\weights(\vx)\right]^\top \in \R^{1 \times P}$.
The $ij$\textsuperscript{th} entry of the matrix $\MKxx \in \R^{N \times N}$ is $\kappa(\vx_i,\vx_j)$, $\vk_{\vx i}$ denotes a vector where
each $j$\textsuperscript{th} element is $\kappa(\vx_i, \vx_j)$, $k_{ii} = \kappa(\vx_i, \vx_i)$ and $\bm\Lambda = - \nabla^2_{\vf \vf}\log p(\vy \mid \vf)$ can be interpreted as per-input noise.
% For notational conciseness, we % \cref{eq-laplace-approx-function-space},
% restrict our notatin to a single function output. The extension to multiple outputs is straightforward.

The GP formulation from \citet{immer2021improving} (shown in \cref{eq-gp-pred-immer})
requires inverting an $N\times N$ matrix which has complexity $O(N^3)$. This limits its applicability to large data sets, which are common in deep learning.

@@

\textbf{Sparse Gaussian processes}
Sparse GPs reduce the computational complexity by representing the GP as a low-rank approximation at a set of inducing
inputs $\vz \in \R^{D}$ with corresponding inducing variables $\vu = f(\vz)$ \citep[see][for an early overview]{quinonero2005unifying}.
The approach by \citet{titsias2009variational} \citep[also used in the DTC approximation, see][]{quinonero2005unifying},
defines the marginal predictive distribution as $q_{\vu}(f_i)  = \int p(f_i  \mid \vu) \, q(\vu) \, \mathrm{d}\vu$.
This give the sparse GP predictive posterior,
\begin{subequations}  \label{eq-svgp-pred}
\begin{align}
  \myexpect_{q_{\vu}(f_i)}[f_i] &= f_{\vw^{*}}(\vx_{i}) + \vk_{\vz i}^{\top}\MKzz^{-1}(\vm - f_{\vw^{*}}(\vz)), \\
  \mathrm{Var}_{q_{\vu}(f_i)}[f_i] &= k_{ii} - \vk_{\vx i}^\top \MKzz^{-1}( \MKzz - \MS_{\vz\vz} )^{-1} \MKzz^{-1} \vk_{\vx i},  \nonumber
\end{align}
\end{subequations}
where the variational distribution is parameterized as $q(\vu) = \mathcal{N}\left(\vu \mid \vm, \MS \right)$ (as in \citet{titsias2009variational,hensman2013gaussian}).
\todo{is this what titsias used or just hensman?}
To the best of our knowledge, there are no methods that convert NNs to sparse GPs.
It is worth noting that the parameters of the variational posterior ($\vm$ and $\MS$) are usually obtained via variational inference.
However, this requires further optimization and is non trivial in the NN setting.

Not only does the GP formulation from \citet{immer2021improving} (shown in \cref{eq-gp-pred-immer}) struggle to scale to large data sets but
it also cannot be incorporate new data by conditioning on it.
This is because it's posterior mean is the NN  $f_{\vw^*}(\vx)$
Our method overcomes both of these limitations via a dual parameterization.
Importantly, our dual parameterization enables us to sparsify it whilst making predictions directly from the GP.

@@

% \newpage
% \section{\our: Sparse function-space representation of neural networks}\label{sec:sfr}
\section{\our: Sparse function-space representation of NNs}\label{sec:sfr}
%\paragraph{GP in the dual parameters}
In this section, we present our method, named \our, which converts a trained NN into a GP.
\our is built upon a dual parameterization of the GP posterior, i.e. an alternative parameterization.
Early work by \citet{csato2002sparse} showed that Expectation Propagation gives rise to a dual parameterization.
More recent work by \citet{adam2021dual,chang2023memory} showed a different dual parameterization arising from the evidence lower bound for sparse variational GPs \citep{hensman2013gaussian,titsias2009variational}.
The dual parameterization from \citet{adam2021dual,chang2023memory}, consists of parameters $\valpha$ and $\vbeta$, which gives rise to the predictive posterior,
% , we can build on the insight from \citet{csato2002sparse} to obtain a `dual' parameterization, $\valpha$ and $\vbeta$,
% The seminal work by \citet{csato2002sparse} (parameterization Lemma~1) gives a parameterization for the posterior process that can be found through the Bayesian update using the GP prior and the likelihood function. This gives a `dual' parameterization, $\valpha$ and $\vbeta$,
% %
\begin{subequations}  \label{eq:gp_pred}
\begin{align}
  \myexpect_{p(f_i \mid\vy)}[f_i] &= \vk_{\vx i}^\top \valpha, \\
  \mathrm{Var}_{p(f_i \mid \vy)}[f_i] &= k_{ii} - \vk_{\vx i}^\top ( \MKxx + \diag(\vbeta)^{-1})^{-1} \vk_{\vx i},  \nonumber
\end{align}
\end{subequations}
%
where $f_i = f_\vw(\vx_i)$ is the function output at input $\vx_i$,
the $ij$\textsuperscript{th} entry of the matrix $\MKxx \in \R^{N \times N}$ is $\kappa(\vx_i,\vx_j)$, $\vk_{\vx i}$ denotes a vector where each $j$\textsuperscript{th} element is $\kappa(\vx_i, \vx_j)$, and $k_{ii} = \kappa(\vx_i, \vx_i)$.
\cref{eq:gp_pred} states that the first two moments of the resultant posterior process (which may not be a GP), can be parameterized via the dual
parameters $\valpha, \vbeta \in \R^{N}$,
% which are the vectors
defined as, \looseness-1
%
\begin{subequations}
\label{eq:dual_param}
\begin{align}
  \alpha_i &\coloneqq \myexpect_{p(\vw \mid \vy)}[\nabla_{f}\log p(y_i \mid f) |_{f=f_i}], \\
  %\quad \text{and} \quad
  \beta_i &\coloneqq - \myexpect_{p(\vw \mid \vy)}[\nabla^2_{f f}\log p(y_i \mid f_i) |_{f=f_i}].
  \end{align}
\end{subequations}
%
The relationships specified are valid for generic likelihoods and involve no approximations since the expectation is under the exact posterior, but given that the model can be expressed in a kernel formulation. \cref{eq:gp_pred} and \cref{eq:dual_param} highlight that the approximate inference technique, usually viewed as a posterior approximation, can be alternatively interpreted as an approximation of the expectation of gradients of loss/likelihoods.

@@

\paragraph{Dual parameters from NN}
Given that we use a Laplace approximation of the NN, we remove the expectation over the posterior \citep[see Ch.~3.4.1 in][for derivation]{rasmussen2006gaussian} and our dual parameters are given by,
%
\begin{subequations}
\label{eq:dual_param_laplace}
\begin{align}
  \hat{\alpha}_i &\coloneqq \nabla_{f}\log p(y_i \mid f) |_{f=f_i} , \\
  %\quad \text{and} \quad
  \hat{\beta}_i &\coloneqq - \nabla^2_{ff}\log p(y_i \mid f) |_{f=f_i}.
  \end{align}
\end{subequations}
%
Substituting \cref{eq:dual_param_laplace} into \cref{eq:gp_pred}, we obtain our GP based on the converged NN. The problem with forming predictions using \cref{eq:gp_pred} is that they cost $O(N^3)$, which limits a full GP model to application on large data sets.

@@

\paragraph{Sparsification via dual parameters}
\label{sec:sparse-dual-gp}
%
Sparse GPs reduce the computational complexity by representing the GP as a low-rank approximation induced by a sparse set of input points \citep[see][for an early overview]{quinonero2005unifying}. Given that we have computed the dual parameters derived from our NN predictions and a kernel function, we could essentially employ any of these sparsification methods. In this work, we opt for the approach suggested by \citet{titsias2009variational} \citep[also used in the DTC approximation, see][]{quinonero2005unifying}, which defines the marginal predictive distribution as $q_{\vu}(f_i)  = \int p(f_i  \mid \vu) \, q(\vu) \, \mathrm{d}\vu$.
Given that we have $p(f_i \mid \vu)$ determined by our GP prior, the goal is to find a $q(\vu)$.
Instead of parameterizing the sparse GP posterior as $q(\vu) = \mathcal{N}\left(\vu \mid \vm, \MS \right)$ (as in \citet{titsias2009variational,hensman2013gaussian})
we follow insights from \citet{adam2021dual} that the posterior under this model bears a structure akin to \cref{eq:gp_pred}.
As such, we project the dual parameters onto the inducing points giving sparse dual parameters,
Using this sparse definition of the dual parameters, our sparse GP posterior takes the following form:
\begin{subequations} \label{eq:dual_sparse_post}
\begin{align}
   \myexpect_{q_{\vu}(\vf)}[f_i] &= \vkzs^{\T} \MKzz^{-1} \valpha_{\vu}, \\
   %\quad \text{and} \quad
   \textrm{Var}_{q_{\vu}(\vf)}[f_i]  &= k_{ii} - \vkzs^\top [\MKzz^{-1} - (\MKzz + \MBeta_{\vu})^{-1} ]\vkzs, \nonumber
\end{align}
\end{subequations}
where $\MKzz$ and $\vkzs$ are defined similarly to $\MKxx$ and $\vk_{\vx i}$ but over the inducing points $\{\vz_j\}_{i=1}^M$, $\vz \in \R^{D}$, instead of the full data set $\dataset$. The key quantities we need to make predictions from our sparse GP from the converged NN are $(\hat{\alpha}_i, \hat{\beta}_i)$ (\cref{eq:dual_param_laplace}) and a kernel function $\kappa$ (\cref{eq-laplace-approx-function-space}). Contrasting \cref{eq:dual_sparse_post} and \cref{eq:gp_pred}, we can see that the computational complexity went from $\mathcal{O}(N^3)$ to $\mathcal{O}(M^3)$, with $M \ll N$.  Crucially, given the structure of our probabilistic model, our sparse dual parameters \cref{eq:dual_sparse} are a compact representation of the full model projected using the kernel.

% As demonstrated in \citet{adam2021dual}, the posterior under this model bears a structure akin to \cref{eq:gp_pred}.
% The authors of that paper exploit the dual parameters for approximate inference but write them using the natural parameterization, primarily because this form is more suitable for optimization through natural gradients.
As such, we project the dual parameters onto the inducing points giving sparse dual parameters,
% In order to link the dual parameters defined in \cref{eq:dual_param}, we write them in the sparse GP using the dual parameters,
%
\begin{equation} \textstyle
  \valpha_{\vu}  =  \sum_{i=1}^N  \vkzi \, \hat{\alpha}_{i}
  \quad \text{and} \quad
  \MBeta_{\vu} =  \sum_{i=1}^N \vkzi \,\hat{\beta}_{i} \, \vkzi^{\T} ,
\label{eq:dual_sparse}
\end{equation}
%
where the sparse dual parameters are now a sum over \emph{all data points}, with $\valpha_{\vu} \in \R^{M}$ and $\MBeta_{\vu} \in \R^{M  \times M}$.

Using this sparse definition of the dual parameters, our sparse GP posterior takes the following form:
\begin{subequations} \label{eq:dual_sparse_post}
\begin{align}
   \myexpect_{q_{\vu}(\vf)}[f_i] &= \vkzs^{\T} \MKzz^{-1} \valpha_{\vu}, \\
   %\quad \text{and} \quad
   \textrm{Var}_{q_{\vu}(\vf)}[f_i]  &= k_{ii} - \vkzs^\top [\MKzz^{-1} - (\MKzz + \MBeta_{\vu})^{-1} ]\vkzs, \nonumber
\end{align}
\end{subequations}
where $\MKzz$ and $\vkzs$ are defined similarly to $\MKxx$ and $\vk_{\vx i}$ but over the inducing points $\{\vz_j\}_{i=1}^M$, $\vz \in \R^{D}$, instead of the full data set $\dataset$. The key quantities we need to make predictions from our sparse GP from the converged NN are $(\hat{\alpha}_i, \hat{\beta}_i)$ (\cref{eq:dual_param_laplace}) and a kernel function $\kappa$ (\cref{eq-laplace-approx-function-space}). Contrasting \cref{eq:dual_sparse_post} and \cref{eq:gp_pred}, we can see that the computational complexity went from $\mathcal{O}(N^3)$ to $\mathcal{O}(M^3)$, with $M \ll N$.  Crucially, given the structure of our probabilistic model, our sparse dual parameters \cref{eq:dual_sparse} are a compact representation of the full model projected using the kernel.

@@

%
We evaluate \our on eight UCI~\citep{UCI} binary and multi-class classification tasks.
Following \citet{immer2021improving}, we used a two-layer MLP with width 50 and tanh activation functions and we used half of the test data set as a validation data set.
We trained the NN using Adam \cite{adam} with a learning rate of $10^{-4}$ and a batch size of $128$.
Training was stopped when the validation NLPD stopped decreasing after 1000 steps.
The checkpoint with lowest NLPD was the NN MAP.
Each experiment was run for $5$ seeds.

>>>>

% Short section names etc
% This must be imported last!
%\usepackage{cleveref}
\usepackage[capitalise,nameinlink]{cleveref}
\crefname{section}{Sec.}{Secs.}
\crefname{algorithm}{Alg.}{Algs.}
\crefname{appendix}{App.}{Apps.}
\crefname{definition}{Def.}{Defs.}
\crefname{table}{Tab.}{Tabs.}

@@

We consider supervised learning with inputs $\vx_i \in \inputDomain$ and outputs $\vy_i \in \outputDomain$ (e.g. regression) or $\vy_{i} \in \{0,1\}^{C}$ (e.g. classification),
giving a data set $\dataset = \{(\vx_{i} , \vy_{i})\}_{i=1}^{N}$.
We introduce a  NN $f_\mathbf{w}: \inputDomain \to \outputDomain$ with weights $\weights \in \R^{P}$ and use a likelihood function $p(\vy \mid f_\mathbf{w}(\vx))$
to link the function values to the output $\vy$ (e.g. Categorical for classification).
For notational conciseness we stick to scalar outputs $y_{i}$.

@@

\textbf{Laplace-GGN}
The Laplace approximation \citet{aa} builds upon this and approximates the weight posterior $p(\vw \mid \mathcal{D}) \approx q(\vw)= \mathcal{N}(\vw^{*} , \bm\Sigma)$
around the MAP weights ($\vw^{*}$) by setting the covariance to the Hessian of the posterior,
\begin{align} \label{eq:laplace-cov}
 \bm\Sigma = - \left[\nabla^{2}_{\vw\vw} \log p(\vw \mid \mathcal{D})  |_{\vw=\vw^{*}} \right]^{-1}.
\end{align}
Computing \cref{eq:laplace-cov} requires calculating the Hessian of the log-likelihood $\nabla^{2}_{\vw\vw} \log p(\vy \mid f_{\vw}(\vx))$ from \cref{eq-empirical-risk}.
As highlighted in \citet{immer2021improving}, computing this Hessian is often infeasible and in practice it is common to adopt the
generalised Gauss-Newton (GGN) approximation,
\begin{align}
 \nabla^{2}_{\vw\vw} \log p(\vy \mid f_{\vw}(\vx)) \approx \Jac{\vw}{\vx}^{\top} \nabla_{\vf\vf}^{2}\log(\vy\mid\vf) \Jac{\vw}{\vx}.
  \nonumber
\end{align}
% \textbf{Laplace-GGN converts BNN into GLM}
\citet{immer2021improving} highlighted that the GGN approximation corresponds to a local linearization of the NN,
% Recent work \citep{khan2019approximate,maddox2021fast} has converted NNs to GPs by linearizing around the NN's MAP weights (obtained from \cref{eq-empirical-risk}).
% Recent work \citep{khan2019approximate,maddox2021fast} has shown that NN (in weight space) leads to a GP, i.e. a function-space approximation.
$f_{\weights^{*}}^{\text{lin}}(\vx) = f_{\weights}(\vx) + \Jac{\weights^{*}}{\vx}(\weights-\weights^{*})$.
This suggests that predictions should be made with a generalised linear model (GLM), given by,
\begin{align} \label{eq-glm}
  p_{\text{\sc GLM}}(\vy \mid \vx, \mathcal{D}) &= \E_{q(\vw)} \left[ p(\vy \mid f_{\vw^{*}}^{\text{lin}}(\vx))\right]. %\nonumber \\
  % &\approx \frac{1}{S} \sum_{s} p(\vy \mid f_{\vw_{s}}^{\text{lin}}(\vx)), \quad \vw_{s} \sim q(\vw)
\end{align}

\textbf{Weight space GLM is equivalent to a GP}
As Gaussian distributions remain tractable under linear transformations, we can covert our model from
weight space to function space \citep[see Ch.~2.1 in ][]{rasmussen2006gaussian}.
 % highlighted that the GGN approximation corresponds to a local linearization of the NN,
As such (and shown in \citet{immer2021improving}), the Bayesian GLM in \cref{eq-glm} has an equivalent GP formulation,
% \begin{align}
%   p_{\text{\sc GP}}(\vy \mid \vx, \mathcal{D}) &= \E_{q(\vf)} \left[ p(\vy \mid \vf)\right] %\nonumber \\
%   % &\approx \frac{1}{S} \sum_{s} p(\vy \mid f_{\vw_{s}}^{\text{lin}}(\vx)), \quad \vw_{s} \sim q(\vw)
% \end{align}
% where the GP posterior is given by,
\begin{subequations}  \label{eq-gp-pred-immer}
\begin{align}
  p_{\text{\sc GP}}(\vy \mid \vx, \mathcal{D}) &= \E_{q(\vf)} \left[ p(\vy \mid \vf)\right] \\ %\nonumber \\
  \myexpect_{p(f_i \mid\vy)}[f_i] &= f_{\vw^{*}}(\vx_{i}), \\
  \mathrm{Var}_{p(f_i \mid \vy)}[f_i] &= k_{ii} - \vk_{\vx i}^\top ( \MKxx + \bm\Lambda^{-1})^{-1} \vk_{\vx i},
\end{align}
\end{subequations}
where the kernel $\kappa(\vx,\vx')=\frac{1}{\delta} \Jac{\weights^*}{\vx} \, \JacT{\weights^*}{\vx'}$ is the Neural Tangent Kernel \citep[NTK,][]{jacot2018neural},
based on the Jacobian at the MAP $\Jac{\weights}{\vx} \coloneqq \left[ \nabla_\weights f_\weights(\vx)\right]^\top \in \R^{1 \times P}$
and $f_i = f_\vw(\vx_i)$ denotes the function output at input $\vx_i$.
The $ij$\textsuperscript{th} entry of the matrix $\MKxx \in \R^{N \times N}$ is $\kappa(\vx_i,\vx_j)$, $\vk_{\vx i}$ denotes a vector where
each $j$\textsuperscript{th} element is $\kappa(\vx_i, \vx_j)$, $k_{ii} = \kappa(\vx_i, \vx_i)$ and $\bm\Lambda = - \nabla^2_{\vf \vf}\log p(\vy \mid \vf)$ can be interpreted as per-input noise.
% For notational conciseness, we % \cref{eq-laplace-approx-function-space},
% restrict our notatin to a single function output. The extension to multiple outputs is straightforward.

@@

\textbf{Sparse Gaussian processes}
The GP formulation from \citet{immer2021improving} (shown in \cref{eq-gp-pred-immer})
requires inverting an $N\times N$ matrix which has complexity $O(N^3)$.
This limits its applicability to large data sets, which are common in deep learning.
Sparse GPs reduce the computational complexity by representing the GP as a low-rank approximation at a set of inducing
inputs $\vz \in \R^{D}$ with corresponding inducing variables $\vu = f(\vz)$ \citep[see][for an early overview]{quinonero2005unifying}.
The approach by \citet{titsias2009variational} \citep[also used in the DTC approximation, see][]{quinonero2005unifying},
defines the marginal predictive distribution as $q_{\vu}(f_i)  = \int p(f_i  \mid \vu) \, q(\vu) \, \mathrm{d}\vu$ where
the variational distribution is parameterized as $q(\vu) = \mathcal{N}\left(\vu \mid \vm, \MS \right)$ (as in \citet{titsias2009variational,hensman2013gaussian}).
\todo{is this what titsias used or just hensman?}
This gives the sparse GP predictive posterior,
\begin{subequations}  \label{eq-svgp-pred}
\begin{align}
  \myexpect_{q_{\vu}(f_i)}[f_i] &= f_{\vw^{*}}(\vx_{i}) + \vk_{\vz i}^{\top}\MKzz^{-1}(\vm - f_{\vw^{*}}(\vz)), \\
  \mathrm{Var}_{q_{\vu}(f_i)}[f_i] &= k_{ii} - \vk_{\vx i}^\top \MKzz^{-1}( \MKzz - \MS )^{-1} \MKzz^{-1} \vk_{\vx i},  \nonumber
\end{align}
\end{subequations}
where $\MKzz$ and $\vkzs$ are defined similarly to $\MKxx$ and $\vk_{\vx i}$ but over the inducing points $\{\vz_j\}_{i=1}^M=\vz \in \R^{M \times D}$.%, instead of the full data set $\dataset$.
% To the best of our knowledge, there are no methods that convert NNs to sparse GPs.
It is worth noting that the parameters of the variational posterior ($\vm$ and $\MS$) are usually obtained via variational
inference, which requires further optimization.% and is non trivial in the NN setting.

Not only does the GP formulation from \citet{immer2021improving} (shown in \cref{eq-gp-pred-immer}) struggle to scale to large data sets but
it also cannot be incorporate new data by conditioning on it.
This is because it's posterior mean is the NN  $f_{\vw^*}(\vx)$.
Our method overcomes both of these limitations via a dual parameterization.
Importantly, our dual parameterization enables us to 1) sparsify the GP withput resorting to further optimisation, and 2)
incorporate new data without retraining (in a computationally efficient manner).

% on new data  make predictions directly from the GP.

@@

% \newpage
% \section{\our: Sparse function-space representation of neural networks}\label{sec:sfr}
\section{\our: Sparse function-space representation of NNs}\label{sec:sfr}
%\paragraph{GP in the dual parameters}
In this section, we present our method, named \our, which converts a trained NN into a GP.
\our is built upon a dual parameterization of the GP posterior.
Early work by \citet{csato2002sparse} showed that Expectation Propagation gives rise to a dual parameterization.
More recent work by \citet{adam2021dual,chang2023memory} showed a different dual parameterization arising from the evidence lower bound for sparse variational GPs \citep{hensman2013gaussian,titsias2009variational}.
The dual parameterization from \citet{adam2021dual,chang2023memory}, consists of parameters $\valpha$ and $\vbeta$, which gives rise to the predictive posterior,
% , we can build on the insight from \citet{csato2002sparse} to obtain a `dual' parameterization, $\valpha$ and $\vbeta$,
% The seminal work by \citet{csato2002sparse} (parameterization Lemma~1) gives a parameterization for the posterior process that can be found through the Bayesian update using the GP prior and the likelihood function. This gives a `dual' parameterization, $\valpha$ and $\vbeta$,
% %
\begin{subequations}  \label{eq:gp_pred}
\begin{align}
  \myexpect_{p(f_i \mid\vy)}[f_i] &= \vk_{\vx i}^\top \valpha, \\
  \mathrm{Var}_{p(f_i \mid \vy)}[f_i] &= k_{ii} - \vk_{\vx i}^\top ( \MKxx + \diag(\vbeta)^{-1})^{-1} \vk_{\vx i},  \nonumber
\end{align}
\end{subequations}
%
% where $f_i = f_\vw(\vx_i)$ is the function output at input $\vx_i$.
% the $ij$\textsuperscript{th} entry of the matrix $\MKxx \in \R^{N \times N}$ is $\kappa(\vx_i,\vx_j)$, $\vk_{\vx i}$ denotes a vector where each $j$\textsuperscript{th} element is $\kappa(\vx_i, \vx_j)$, and $k_{ii} = \kappa(\vx_i, \vx_i)$.
\cref{eq:gp_pred} states that the first two moments of the resultant posterior process (which may not be a GP), can be parameterized via the dual
parameters $\valpha, \vbeta \in \R^{N}$,
% which are the vectors
defined as, \looseness-1
%
\begin{subequations}
\label{eq:dual_param}
\begin{align}
  \alpha_i &\coloneqq \myexpect_{p(\vw \mid \vy)}[\nabla_{f}\log p(y_i \mid f) |_{f=f_i}], \\
  %\quad \text{and} \quad
  \beta_i &\coloneqq - \myexpect_{p(\vw \mid \vy)}[\nabla^2_{f f}\log p(y_i \mid f_i) |_{f=f_i}].
  \end{align}
\end{subequations}
%
The relationships specified are valid for generic likelihoods and involve no approximations since the expectation is under the
exact posterior, given that the model can be expressed in a kernel formulation.
\cref{eq:gp_pred} and \cref{eq:dual_param} highlight that the approximate inference technique, usually viewed as a posterior approximation,
can alternatively be interpreted as an approximation of the expectation of loss (likelihood) gradients.

@@

\paragraph{Dual parameters from NN}
Given that we use a Laplace approximation of the NN, we remove the expectation over the posterior \citep[see Ch.3.4.1 in][for derivation]{rasmussen2006gaussian} and our dual parameters are given by,
%
\begin{subequations}
\label{eq:dual_param_laplace}
\begin{align}
  \hat{\alpha}_i &\coloneqq \nabla_{f}\log p(y_i \mid f) |_{f=f_i} , \\
  %\quad \text{and} \quad
  \hat{\beta}_i &\coloneqq - \nabla^2_{ff}\log p(y_i \mid f) |_{f=f_i}.
  \end{align}
\end{subequations}
%
Substituting \cref{eq:dual_param_laplace} into \cref{eq:gp_pred}, we obtain our GP based on the converged NN. The problem with forming predictions using \cref{eq:gp_pred} is that they cost $O(N^3)$, which limits its applicability to large data sets.

@@

\paragraph{Sparsification via dual parameters}
\label{sec:sparse-dual-gp}
%
% Sparse GPs reduce the computational complexity by representing the GP as a low-rank approximation induced by a sparse set of input points \citep[see][for an early overview]{quinonero2005unifying}.
Given that we have computed the dual parameters derived from our NN predictions and a kernel function, we could essentially employ any sparsification method \citep{quinonero2005unifying}.
In this work, we opt for the approach suggested by \citet{titsias2009variational} \citep[also used in the DTC approximation, see][]{quinonero2005unifying}, which defines the marginal predictive distribution as $q_{\vu}(f_i)  = \int p(f_i  \mid \vu) \, q(\vu) \, \mathrm{d}\vu$.
% Given that we have $p(f_i \mid \vu)$ determined by our GP prior, the goal is to find a $q(\vu)$.
Instead of parameterizing the sparse GP posterior as $q(\vu) = \mathcal{N}\left(\vu \mid \vm, \MS \right)$ (as in \citet{titsias2009variational,hensman2013gaussian})
we follow insights from \citet{adam2021dual} that the posterior under this model bears a structure akin to \cref{eq:gp_pred}.
As such, we project the dual parameters onto the inducing points giving sparse dual parameters,
Using this sparse definition of the dual parameters, our sparse GP posterior takes the following form:
\begin{subequations} \label{eq:dual_sparse_post}
\begin{align}
   \myexpect_{q_{\vu}(\vf)}[f_i] &= \vkzs^{\T} \MKzz^{-1} \valpha_{\vu}, \\
   %\quad \text{and} \quad
   \textrm{Var}_{q_{\vu}(\vf)}[f_i]  &= k_{ii} - \vkzs^\top [\MKzz^{-1} - (\MKzz + \MBeta_{\vu})^{-1} ]\vkzs, \nonumber
\end{align}
\end{subequations}
% where the key quantities we need to make predictions from our sparse GP from the converged NN are $(\hat{\alpha}_i, \hat{\beta}_i)$ (\cref{eq:dual_param_laplace}) and a kernel function $\kappa$ (\cref{eq-laplace-approx-function-space}).
% As demonstrated in \citet{adam2021dual}, the posterior under this model bears a structure akin to \cref{eq:gp_pred}.
% The authors of that paper exploit the dual parameters for approximate inference but write them using the natural parameterization, primarily because this form is more suitable for optimization through natural gradients.
where the dual parameters are projected onto the inducing points to get our sparse dual parameters,
% In order to link the dual parameters defined in \cref{eq:dual_param}, we write them in the sparse GP using the dual parameters,
%
\begin{equation} \textstyle
  \valpha_{\vu}  =  \sum_{i=1}^N  \vkzi \, \hat{\alpha}_{i}
  \quad \text{and} \quad
  \MBeta_{\vu} =  \sum_{i=1}^N \vkzi \,\hat{\beta}_{i} \, \vkzi^{\T} .
\label{eq:dual_sparse}
\end{equation}
%
Note that the sparse dual parameters are now a sum over \emph{all data points}, with $\valpha_{\vu} \in \R^{M}$ and $\MBeta_{\vu} \in \R^{M  \times M}$.
Contrasting \cref{eq:dual_sparse_post} and \cref{eq:gp_pred}, we can see that the computational complexity went from $\mathcal{O}(N^3)$ to $\mathcal{O}(M^3)$, with $M \ll N$.
Crucially, given the structure of our probabilistic model, our sparse dual parameters \cref{eq:dual_sparse} are a compact representation of the full model projected using the kernel.

% Using this sparse definition of the dual parameters, our sparse GP posterior takes the following form:
% \begin{subequations} \label{eq:dual_sparse_post}
% \begin{align}
%    \myexpect_{q_{\vu}(\vf)}[f_i] &= \vkzs^{\T} \MKzz^{-1} \valpha_{\vu}, \\
%    %\quad \text{and} \quad
%    \textrm{Var}_{q_{\vu}(\vf)}[f_i]  &= k_{ii} - \vkzs^\top [\MKzz^{-1} - (\MKzz + \MBeta_{\vu})^{-1} ]\vkzs, \nonumber
% \end{align}
% \end{subequations}
% where $\MKzz$ and $\vkzs$ are defined similarly to $\MKxx$ and $\vk_{\vx i}$ but over the inducing points $\{\vz_j\}_{i=1}^M$, $\vz \in \R^{D}$, instead of the full data set $\dataset$. The key quantities we need to make predictions from our sparse GP from the converged NN are $(\hat{\alpha}_i, \hat{\beta}_i)$ (\cref{eq:dual_param_laplace}) and a kernel function $\kappa$ (\cref{eq-laplace-approx-function-space}). Contrasting \cref{eq:dual_sparse_post} and \cref{eq:gp_pred}, we can see that the computational complexity went from $\mathcal{O}(N^3)$ to $\mathcal{O}(M^3)$, with $M \ll N$.  Crucially, given the structure of our probabilistic model, our sparse dual parameters \cref{eq:dual_sparse} are a compact representation of the full model projected using the kernel.

@@

%
We evaluate \our on eight UCI~\citep{UCI} binary and multi-class classification tasks.
Following \citet{immer2021improving}, we used a two-layer MLP with width 50 and tanh activation functions and we used half of the test data set as a validation data set.
We trained the NN using Adam \cite{adam} with a learning rate of $10^{-4}$ and a batch size of $128$.
Training was stopped when the validation NLPD stopped decreasing after 1000 steps.
The checkpoint with lowest NLPD was the NN MAP.
Each experiment was run for $5$ seeds.\looseness-1
